Commutative Character Sheaves

### 2015-11-14 followup.tex
Fixed a few typos

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ is a morphism of $H\times H$-equivariant Weil local systems on $G\times G$; likewise $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$. 
\end{lemma}

@@

Now suppose $q : G \to G_0$ is a regular epimorphism of smooth group schemes with kernel pair $(a,p)$
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G_0.
\end{tikzcd}
\]
Let $\cs{L}_0$ be a linear character sheaf on $G_0$.
Consider the functor
\[
q^* : \CS(G_0) \to \CS(G)
\]
given on objects by $(\cs{L}_0,\mu_0) \mapsto (q^*\cs{L}_0, (q\times q)^* \mu_0)$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
%To see that $(q^*\cs{L}_0, (q\times q)^* \mu_0)$ is indeed a linear character sheaf on $G$, verify \cite{cunningham-roe:13a}, CS.3], arguing as in \cite{cunningham-roe:13a}].
This linear character sheaf $(\cs{L},\mu)\ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$ on $G$ comes equipped with the canonical isomorphism $\nu : a^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of canonical isomorphisms.
\[
\begin{tikzcd}[column sep=40]
a^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
a^* ( q^*\cs{L}_0) \arrow{r}{q\circ a = q\circ p}
& p^* (q^*\cs{L}_0)
\end{tikzcd}
\]
Since this isomorphism satisfies \eqref{ECS1}, it follows that $(\cs{L}, \mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$.
If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(G_0)$, 
then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, arguing as in \cite{cunningham-roe:13a} so $\alpha$ is a morphism in $\CS(G)$.
%\todo{I find myself having to make many references to \cite{cunningham-roe:13a}] and promising that the results there adapt to the noncommutative group case. It makes me uneasy, but the alternative is to actually repeat the proofs here, with the weaker hypothesis on $G$, which would be ugly. I'm just agonizing out loud, I guess.} 
These simple observations define the comparison functor
\[
q_H^* : \CS(G_0) \to CS_H(G)
\]
and show that the functor $q^* : \CS(G_0) \to \CS(G)$ factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(G_0) \arrow{dl}{q_H^*}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]
The definition of $q_H^* : \CS(G_0)\to \CS_H(G)$ will be revisited in the proof of the following result.

@@

With this description of $q_H^*: \CS(Q)\to \CS_H(G)$ is is easy to see that it is an equivalence.
Its adjoint is given on objects by $(\cs{L},\mu,\nu) \mapsto (R_1(\cs{L},\nu), R_2(\nu))$, making use of Lemma~\ref{lem:HH}, and on morphisms by $\alpha \mapsto R_1(\alpha)$.
The adjunction is built from the adjunctions for $(L_1,R_1)$ and $(L_2, R_2)$.
\end{proof}

\subsection{Closed subgroups}\label{ssec:closed}

We now suppose that $j : H \hookrightarrow G$ is a closed subgroup and that the action $a : H\times G\to G$ is obtained by restricting the action $m : G \times G\to G$ to $H\times G$.
%Then we have the diagram
%\[
%\begin{tikzcd}
%\ &  \arrow{dl}[swap]{a} H \times G \arrow{dr}{p} \arrow{dd}[swap]{f} & \\
%G & & G \\
%& \arrow{ul}{m} G \times G \arrow{ur}[swap]{p_2} & 
%\end{tikzcd}
%\]
Define $f : H\times G \to G\times G$ by $f(h,g) = (j(h),g)$ and note that $a = m\circ f$.
Every $H$-equivariant linear character sheaf $(\cs{L},\mu, \nu)$ uniquely determines an isomorphism 
\[
\beta : \cs{L}\vert_H \to (\EE)_{H}
\]
by 
\begin{equation}\label{betanu}
\nu = (\beta \boxtimes \id) \circ  f^*\mu,
\end{equation}
or more precisely, by the following diagram of isomorphisms, where we write $p_1: G\times G\to G$ for projection to the first component and $p_2 : G\times G \to G$ for projection to the second.
\[
\begin{tikzcd}[column sep=30]
\ & a^*\cs{L} \arrow{dl}{m\circ f = a} \arrow{r}{\nu} & p^*\cs{L} & \\
f^* m^*\cs{L} \arrow{dr}{f^*\mu} &&& (\EE)_{H} \boxtimes \cs{L} \arrow{ul} \\ 
& f^*(\cs{L}\boxtimes \cs{L}) \arrow{r}{p_1 \circ f = j}[swap]{p_2\circ f = \id}  & 
\cs{L}\vert_{H} \boxtimes \cs{L} \arrow{ur}{\beta\boxtimes \id} & 
\end{tikzcd}
\]
In this way we see we may replace $(\cs{L},\mu,\nu)$ with a triple, $(\cs{L},\mu,\beta)$, from which $(\cs{L},\mu,\nu)$ may be recovered.
The next lemma makes that statement more precise.

\begin{lemma}\label{lem:beta}
$\CS_H(G)$ is equivalent to the category $\CS^H(G)$ of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu)\in \CS(G)$ and $\beta : \cs{L}\vert_H \to (\EE)_H$ is an isomorphism in $\CS(H)$, with morphisms $\alpha : (\cs{L},\mu,\beta) \to (\cs{L}',\mu',\beta')$ defined by the condition
$\alpha : (\cs{L},\mu) \to (\cs{L}',\mu')$ is a morphism in $\CS(G)$ and $\beta = \beta' \circ \alpha\vert_{H}$.
\end{lemma}

>>>>

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ and $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are morphisms of $H\times H$-equivariant Weil local systems on $G\times G$. 
\end{lemma}

@@

Now suppose $q : G \to G_0$ is a regular epimorphism of smooth group schemes with kernel pair $(a,p)$
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G_0.
\end{tikzcd}
\]
Let $\cs{L}_0$ be a linear character sheaf on $G_0$.
Consider the functor
\[
q^* : \CS(G_0) \to \CS(G)
\]
given on objects by $(\cs{L}_0,\mu_0) \mapsto (q^*\cs{L}_0, (q\times q)^* \mu_0)$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
%To see that $(q^*\cs{L}_0, (q\times q)^* \mu_0)$ is indeed a linear character sheaf on $G$, verify \cite{cunningham-roe:13a}, CS.3], arguing as in \cite{cunningham-roe:13a}].
This linear character sheaf $(\cs{L},\mu)\ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$ on $G$ comes equipped with the canonical isomorphism $\nu : a^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of canonical isomorphisms.
\[
\begin{tikzcd}[column sep=40]
a^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
a^* ( q^*\cs{L}_0) \arrow{r}{q\circ a = q\circ p}
& p^* (q^*\cs{L}_0)
\end{tikzcd}
\]
Since this isomorphism satisfies \eqref{ECS1}, it follows that $(\cs{L}, \mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$.
If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(G_0)$, 
then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, so $\alpha$ is a morphism in $\CS(G)$.
%\todo{I find myself having to make many references to \cite{cunningham-roe:13a}] and promising that the results there adapt to the noncommutative group case. It makes me uneasy, but the alternative is to actually repeat the proofs here, with the weaker hypothesis on $G$, which would be ugly. I'm just agonizing out loud, I guess.} 
These simple observations define the comparison functor
\[
q_H^* : \CS(G_0) \to CS_H(G)
\]
and show that the functor $q^* : \CS(G_0) \to \CS(G)$ factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(G_0) \arrow{dl}{q_H^*}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]
The definition of $q_H^* : \CS(G_0)\to \CS_H(G)$ will be revisited in the proof of the following result.

@@

With this description of $q_H^*: \CS(Q)\to \CS_H(G)$ is is easy to see that it is an equivalence.
Its adjoint is given on objects by $(\cs{L},\mu,\nu) \mapsto (R_1(\cs{L},\nu), R_2(\nu))$, making use of Lemma~\ref{lem:HH}, and on morphisms by $\alpha \mapsto R_1(\alpha)$.
The adjunction is built from the adjunctions $(\epsilon_1, \eta_1)$ and $(\epsilon_2, \eta_2)$ for $(L_1,R_1)$ and $(L_2, R_2)$.
\end{proof}

\subsection{Quotient by a closed subgroup}\label{ssec:closed}

We now suppose that $j : H \hookrightarrow G$ is a closed subgroup scheme over $\Fq$ and that the action $a : H\times G\to G$ is obtained by restricting the action $m : G \times G\to G$ to $H\times G$.
%Then we have the diagram
%\[
%\begin{tikzcd}
%\ &  \arrow{dl}[swap]{a} H \times G \arrow{dr}{p} \arrow{dd}[swap]{f} & \\
%G & & G \\
%& \arrow{ul}{m} G \times G \arrow{ur}[swap]{p_2} & 
%\end{tikzcd}
%\]
Define $f : H\times G \to G\times G$ by $f(h,g) = (j(h),g)$ and note that $a = m\circ f$.
Every $H$-equivariant linear character sheaf $(\cs{L},\mu, \nu)$ uniquely determines an isomorphism 
\[
\beta : \cs{L}\vert_H \to (\EE)_{H}
\]
by 
\begin{equation}\label{betanu}
\nu = (\beta \boxtimes \id) \circ  f^*\mu,
\end{equation}
or more precisely, by the following diagram of isomorphisms, where we write $p_1: G\times G\to G$ for projection to the first component and $p_2 : G\times G \to G$ for projection to the second.
\[
\begin{tikzcd}[column sep=30]
\ & a^*\cs{L} \arrow{dl}{m\circ f = a} \arrow{r}{\nu} & p^*\cs{L} & \\
f^* m^*\cs{L} \arrow{dr}{f^*\mu} &&& (\EE)_{H} \boxtimes \cs{L} \arrow{ul} \\ 
& f^*(\cs{L}\boxtimes \cs{L}) \arrow{r}{p_1 \circ f = j}[swap]{p_2\circ f = \id}  & 
\cs{L}\vert_{H} \boxtimes \cs{L} \arrow{ur}{\beta\boxtimes \id} & 
\end{tikzcd}
\]
In this way we see we may replace $(\cs{L},\mu,\nu)$ with a triple, $(\cs{L},\mu,\beta)$, from which $(\cs{L},\mu,\nu)$ may be recovered.
The next lemma makes that statement more precise.

\begin{lemma}\label{lem:beta}
If $H\hookrightarrow G$ is a closed subgroup scheme over $\Fq$ then
$\CS_H(G)$ is equivalent to the category $\CS^H(G)$ of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu)\in \CS(G)$ and $\beta : \cs{L}\vert_H \to (\EE)_H$ is an isomorphism in $\CS(H)$; morphisms $(\cs{L},\mu,\beta) \to (\cs{L}',\mu',\beta')$ in $\CS^H(G)$ are morphisms $\alpha: \cs{L} \cs{L}'$ in $\CS(G)$ such that $\beta = \beta' \circ \alpha\vert_{H}$.
\end{lemma}
